Mutual Information Plots

### 2025-04-21 mutual_information_curves.qmd

    # Loop through each row of prob_df
    for (j in 1:nrow(prob_df)) {
      # Calculate the squared difference between x_i and y_i for each row j
      diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
    w[j] <- w[j] + rank_weights[i]* diff^2  # Sum squared differences for each row
    }
  }
  w <- ifelse(w == 0.0000, 1, 1 / w)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
  w <- w/sum(w)
  # Compute weights dynamically for each row
  weighted_df <- prob_df %>%
    mutate(w = w)

  mutual_info <- factorial(num_teams) * sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/nrow(weighted_df) * weighted_df$probability)))

  # Return the weighted data frame with mutual information weights
  return(mutual_info)
}
mut_info(all_results_norm,8, c(10,5,1,0,0,0,0,0))
mut_info(de_norm_results_eight,8, c(10,5,1,0,0,0,0,0))
mut_info(test_rr_noTies,8, c(10,5,1,0,0,0,0,0))
mut_info(seven_games_se_results,8, c(10,5,1,0,0,0,0,0))

de_norm_results_eight <- de_norm_results %>% filter(teams == 8)
de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)
all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)
seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)
# with randomly assigning ties
test_rr_noTies <- test_rr %>%
  group_by(simulation) %>%
  mutate(rank_hat = rank(rank_hat,ties.method="random"))

@@

# Compute mutual information for each dataset
mi_de_norm <- compute_mut_plots(de_norm_results_eight)
mi_se_norm <- compute_mut_plots(all_results_norm)
mi_test_rr <- compute_mut_plots(test_rr_noTies)
mi_seven_se <- compute_mut_plots(seven_games_se_results)

# Plot results
plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm, mi_test_rr, mi_seven_se)),
     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Mutual Information vs. Number of 1's")

# Add other datasets
lines(indices, mi_se_norm, type="b", pch=19, col="red")
lines(indices, mi_test_rr, type="b", pch=19, col="green")
lines(indices, mi_seven_se, type = "b", pch=19, col="black")

# Add legend
legend("topleft", legend=c("Double Elimination", "Single Elimination", "Round Robin", "7 Game Single Elimination"), 
       col=c("blue", "red", "green", "black"), pch=19, lty=1)

### Single Elimination vs Double Elimination ###

# Plot results
plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm)),
     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Single Elimination vs Double Elimination")

# Add other datasets
lines(indices, mi_se_norm, type="b", pch=19, col="red")

# Add legend
legend("bottomright", legend=c("Double Elimination", "Single Elimination"), 
       col=c("blue", "red"), pch=19, lty=1)

```

>>>>

    # Loop through each row of prob_df
    for (j in 1:nrow(prob_df)) {
      # Calculate the squared difference between x_i and y_i for each row j
      diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
    w[j] <- w[j] + rank_weights[i]* diff^2  # Sum squared differences for each row
    }
  }
  w <- 1 / (w + 1)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
  w <- w/sum(w)
  # Compute weights dynamically for each row
  weighted_df <- prob_df %>%
    mutate(w = w)

  mutual_info <- factorial(num_teams) * sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/factorial(num_teams) * weighted_df$probability)))

  # Return the weighted data frame with mutual information weights
  return(mutual_info)
}
mut_info(all_results_norm,8, c(10,5,1,0,0,0,0,0))
mut_info(de_norm_results_eight,8, c(10,5,1,0,0,0,0,0))
mut_info(test_rr_noTies,8, c(10,5,1,0,0,0,0,0))
mut_info(seven_games_se_results,8, c(10,5,1,0,0,0,0,0))
mut_info(perms_df,8,c(1,0,0,0,0,0,0,0))

de_norm_results_eight <- de_norm_results %>% filter(teams == 8)
de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)
test_unif$simulation <- rep(1:10000, each = nrow(test_unif) / 10000)
equal_strengths$simulation <- rep(1:10000, each = nrow(equal_strengths) / 10000)
all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)
bad_seeding_structure$simulation <- rep(1:10000, each = nrow(bad_seeding_structure) / 10000)
seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)
# with randomly assigning ties
test_rr_noTies <- test_rr %>%
  group_by(simulation) %>%
  mutate(rank_hat = rank(rank_hat,ties.method="random"))

@@

perms_list <- permn(8)
all_data <- list()
# Loop through each permutation and create the data frame
for (i in 1:length(perms_list)) {
  # Create a temporary data frame for the current permutation
  temp_df <- data.frame(
    true_rank = rep(1:8, each = 1),  # Each row gets a true_rank from 1 to 8
    rank_hat = perms_list[[i]],      # The current permutation's ranks
    simulation = rep(i, 8)           # Assign the simulation number to each row
  )
  # Add the temporary data frame to the list
  all_data[[i]] <- temp_df
}
# Combine all the data frames into one
perms_df <- do.call(rbind, all_data)

# Compute mutual information for each dataset
mi_de_norm <- compute_mut_plots(de_norm_results_eight)
mi_se_norm <- compute_mut_plots(all_results_norm)
mi_test_rr <- compute_mut_plots(test_rr_noTies)
mi_seven_se <- compute_mut_plots(seven_games_se_results)
mi_random <- compute_mut_plots(equal_strengths)
mi_se_unif <- compute_mut_plots(test_unif)
mi_bad <- compute_mut_plots(bad_seeding_structure)
mi_perms <- compute_mut_plots(perms_df)

# Plot results
plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm, mi_test_rr, mi_seven_se, mi_random, mi_se_unif)),
     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Mutual Information vs. Number of 1's")

# Add other datasets
lines(indices, mi_se_norm, type="b", pch=19, col="red")
lines(indices, mi_test_rr, type="b", pch=19, col="green")
lines(indices, mi_seven_se, type = "b", pch=19, col="black")
lines(indices, mi_random, type="b", pch=19, col="purple")
lines(indices, mi_se_unif, type="b", pch=19, col="orange")

# Add legend
legend("topleft", legend=c("Double Elimination", "Single Elimination", "Round Robin", "7 Game Single Elimination", "Random Single Elimination", "Uniform Single Elimination"), 
       col=c("blue", "red", "green", "black", "purple", "orange"), pch=19, lty=1)

mi_plots_df <- data.frame(
  indices = indices,
  DE = mi_de_norm,
  Norm_SE = mi_se_norm,
  Random_SE = mi_random,
  Unif_SE = mi_se_unif,
  Bad_Seeds_SE = mi_bad,
  Perms = mi_perms
)
# Reshape data into long format for ggplot
df_long <- pivot_longer(mi_plots_df, cols = -indices, names_to = "Method", values_to = "MI")
ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
  geom_line() +           # Add lines
  geom_point() +          # Add points
  labs(
    title = "Single Elimination vs Double Elimination",
    x = "Number of 1's in the vector",
    y = "Mutual Information Bits"
  ) +
  theme_minimal() +       # Use a clean theme
  theme(legend.position = "right")
```
